\bf\mftitle

### 2011-01-13 doc/jInfer/full/architecture/architecture.tex
minors

For even more generality in design, we decided to implement abstract classes in midlevel:
\begin{itemize}
	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metada}),
	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
\end{itemize}
As practice showed, for methods handling and infering structural properties, it is important to recognize whether structural node on input is element or text node.
However methods for content devising don't need to know, if they are working on infering model for content of attribute or text node.

@@

TODO anti
% hey vektor, what to DO here?

\subsection{Rules, Grammars, etc}

@@

\section{NetBeans Modules}

@@

At the present, jInfer modules use the following metadata, all defined in \code{IGGUtils} class:
\begin{itemize}
	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
	\item \code{is.sentinel}, filled in by  BasicIGG and XSDImport module: when importing XML document, one can build whole tree in memory (as in fig. \ref{xml_example}) and then construct list of rules from it (thus saving memory). With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import), but right side can be defined anywhere in source file. To save complicated loading of whole schema, searching and pairing elements thorough rules, nodes on right side of rule are created empty - holding only name of node. Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
	All of these metadata are of set/not set character (using \code{Boolean.TRUE}).
\end{itemize}

@@

\subsection{Nondeterministic Finite Automaton}
For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.

Let's begin by smallest of the classes, the \code{State<T>}.
It represents automaton state, it has two integer members: \code{name} and \code{finalCount}.
Name clearly serves as name of state in visualization and to string conversion.
Final count is for representing whether the state is final in automaton.
The field is not true/false but integral to help algorithms use stastistics over automatons (how many times in XML input is this state final?).

Next we have \code{Step<T>} which stands for automaton transition.
It has its \code{source} and \code{destination} states references.
Symbol accepted by using this transition is store in \code{acceptsSymbol} member of generic type \code{T}.
And finally member \code{useCount}, which is integer stating how many times the transition was used when constructing prefix tree automaton from input data.
Simplifying algorithm can use this number for statistic purposes.

\code{Automaton<T>} class puts theese together into nondeterministic finite automaton.
It has reference to \code{initialState}, it has \code{newStateName} integer value to assure unique state names inside one automaton (incrementing every time new state creates).
We use two maps to implement transition function (delta($\delta$) function  in most literature).
One map of type \code{Map<State<T>, Set<Step<T>>>} called \code{delta} represents mapping from state into set of all outgoing transitions from state.
Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better perfomance).
There exist only one instance of each step from one state to another.
That instance is referenced in delta map (on place of source state), and in reverse delta map (on place of destination state).
Loops are no speciality, just source = destination.

Automaton supports creating automaton as a copy of another one (not reference copy, but true copy), this can be used when searching solution space to create more versions of automaton to edit.
We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
Create empty automaton and then call this method for every positive example input string of language.
You will get PTA with useCounts and finalCounts set properly on steps/states.

The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates()} method.
Method merges second state given into first one (or all states in list into first one in list).
All \{in | out\}-transitions are redirected properly (or discarted as needed).
Use count and final count are updated as sums of values from merging transitions/states.
One can ask for merging states, that are merged out from automaton already.
Take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
One asks to merge states 3 and 4.
State 3 then becomes final state with loop and state 4 disappears (fig. \ref{automaton_merge2}).
If then one ask to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).

@@

The most important method in each module is \code{start}, defined in each of the interfaces. This method is called by the framework when the respective step of inference is being executed. It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished. We will look at both parameter now in more detail.

@@

\emph{Initial Grammar Generator} takes an object of type \emph{Input}. This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}. Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.\\

@@

% TODO vektor Talk about Capabilities

>>>>

For even more generality in design, we decided to implement abstract classes in midlevel:
\begin{itemize}
	\item \code{AbstractNamedNode}, which implements methods from \code{NamedNode} interface to handle context, name and metadata (will discuss later in section \ref{section_metadata}),
	\item \code{AbstractStructuralNode}, which implements only task of deciding if instance is \code{Element} or \code{SimpleData} actually.
\end{itemize}
As practice showed, for methods handling and infering structural properties, it is important to recognize whether structural node on input is element or text node.
However methods for content devising don't need to know, if they are working on infering model for content of attribute or text node.

@@

\subsection{Rules and grammars}

@@

\subsection{Nondeterministic Finite Automaton}
We recommend skipping this section in first read as it describes advanced features.
For all inference algorithms based on merging states of NFA's, our implementation of nondeterministic finite automaton might be interesting.
Implementation consists of 4 classes: \code{Automaton<T>}, \code{Step<T>}, \code{State<T>} and \code{AutomatonCloner<A, B>} (see package \code{cz.cuni.mff.ksi.jinfer.base.automaton}).
Whole implementation uses java generics for representation of symbol of alphabet, we denote by \code{T} the java type of symbol.

\paragraph{State}
Let's begin by smallest of the classes, the \code{State<T>}.
It represents automaton state, it has two integer members: \code{name} and \code{finalCount}.
Name clearly serves as name of state in visualization and to string conversion.
Final count is for representing whether the state is final in automaton.
The field is not true/false but integral to help algorithms use stastistics over automatons (how many times in XML input is this state final?).

\paragraph{Step}
Next we have \code{Step<T>} which stands for automaton transition.
It has its \code{source} and \code{destination} states references.
Symbol accepted by using this transition is store in \code{acceptsSymbol} member of generic type \code{T}.
And finally member \code{useCount}, which is integer stating how many times the transition was used when constructing prefix tree automaton from input data.
Simplifying algorithm can use this number for statistic purposes.

\paragraph{Automaton}
\code{Automaton<T>} class puts theese together into nondeterministic finite automaton.
It has reference to \code{initialState}, it has \code{newStateName} integer value to assure unique state names inside one automaton (incrementing every time new state is created).
We use two maps to implement transition function ($\delta$-function).
One map of type \code{Map<State<T>, Set<Step<T>{}>{}>} called \code{delta} represents mapping from state into set of all outgoing transitions from state.
Second is just reversed map, called \code{reverseDelta}, which holds all incoming transitions into state (for better perfomance).
There exist only one instance of each step from one state to another.
That instance is referenced in delta map (on place of source state), and in reverse delta map (on place of destination state).
Loops are no speciality, just source = destination.

Automaton supports creating automaton as a copy of another one (not reference copy, but true copy), this can be used when searching solution space to create more versions of automaton to edit.
We implemented building of prefix-tree automaton (PTA) in \code{buildPTAOnSymbol()} method.
Create empty automaton and then call this method for every input string of language.
You will get PTA with useCounts and finalCounts set properly on steps/states.

The biggest thing we offer to scientists is state merging by simply calling \code{mergeStates(state1, state2)} method.
Method merges second state given into first one (or an overloaded version - all states in list into first one in list).
All \{in|out\}-transitions are redirected properly (or discarted as needed).
Variables useCount and finalCount are updated to sums of values from merged transitions/states.
One can ask for merging states, that are merged out from automaton already.
Lets take example automaton of fig. \ref{automaton_merge1} (states are labeled [name|finalCount], steps by \{symbol|useCount\}).
One asks to merge states 3 and 4.
State 3 then becomes final state with loop and state 4 disappears (fig. \ref{automaton_merge2}).
If then one ask to merge states 2 and 4, automaton properly handles situation by knowing, that old state 4, was merged into state 3, and merges states 2 and 3 (see fig. \ref{automaton_merge3}).
This can be useful in $k,h-context$ and $s,k-strings$ implementations (find all context/strings, then supply list of states to merge and don't bother with state names updates).

@@

At the present, jInfer modules use the following metadata, all defined in \code{IGGUtils} class:
\begin{itemize}
	\item \code{from.schema}, filled in by BasicIGG and XSDImport modules: means that this rule was created (originates from) from a schema.
	\item \code{from.query}, filled in by BasicIGG and XSDImport modules: means that this rule was created from a query.
	\item \code{is.sentinel}, filled in by  BasicIGG and XSDImport module: when importing XML document, one can build whole tree in memory (as in fig. \ref{xml_example}) and then construct list of rules from it (thus saving memory). With schema import however, one not only doesn't know right side of rule in advance (solved by stack in XML import), but right side can be defined anywhere in source file. To save complicated loading of whole schema, searching and pairing elements thorough rules, nodes on right side of rule are created empty - holding only name of node. Fact, that this node has no more information than it's name and position on right side of rule is denominated by labeling the node as sentinel.
	\item \code{schema.data}, filled in by XSDImport module: TODO reseto Explain
\end{itemize}
All of these metadata are of set/not set character (using \code{Boolean.TRUE}).

@@

@@

The most important method in each module is \code{start}, defined in each of the interfaces. This method is called by the framework when the respective step of inference is to be executed. It has always two parameters: the actual input data for the module, and a callback object to report to when this step is finished. We will look at both parameters now in more detail.

@@

\jmodule{Initial Grammar Generator} takes an object of type \emph{Input}. This class encapsulates all the input files in 3 collections of \code{File}: \code{documents}, \code{schemas} and \code{queries}. Enumerating these files provides IGG with access to all data it needs to create Initial Grammar.\\

@@

% TODO vektor Talk about Capabilities
\section{NetBeans Modules}
